mathml: Translate R expressions to MathML and LaTeX/MathJax

### 2023-01-10 rj/mathml.Rmd
improve table with custom operators

```{r custom-operators, echo=FALSE}
op1 <- list(
  cross=quote(A %*% B),
  dot=quote(A %.% B),
  kronecker=quote(A %x% B),
  'integer division'=quote(A %/% B),
  modulo=quote(A %% B),
  and=quote(A & B),
  vel=quote(A | B),
  xor=quote(xor(A, B)),
  not=quote(!A),
  equal=quote(A == B),
  assign=quote(A <- B))

@@

op2 <- list(
  different=quote(A != B),
  tilde=quote(A ~ B),
  approx=quote(A %~~% B),
  equiv=quote(A %==% B),
  congr=quote(A %=~% B),
  prop=quote(A %prop% B),
  'in'=quote(A %in% B),
  intersect=quote(intersect(A, B)),
  union=quote(union(A, B)),
  crossprod=quote(crossprod(A, B)),
  intersect=quote(is.null(A)))

@@

op3 <- list(
  twosidearrow=quote(A %<->% B),
  rightarrow=quote(A %->% B),
  leftarrow=quote(A %<-% B),
  uparrow=quote(A %up% B),
  down=quote(A %down% B),
  iff=quote(A %<=>% B),
  rArr=quote(A %=>% B),
  lArr=quote(A %<=% B),
  dblup=quote(A %dblup% B),
  dbldown=quote(A %dbldown% B),
  " "="")

@@

op3 <- as.character(op3)
if(knitr::is_latex_output())
  op3 <- sapply(op3, FUN=knitr:::escape_latex)
if(knitr::is_html_output())
  op3 <- sapply(op3, FUN=knitr:::escape_html)

@@

```{r, results='asis'}
zv <- Vectorize(z, vectorize.args='tau')

@@

The degrees of freedom result from the 13 predicted mean response times,
minus 5 free parameters (\(c\) is a redundant scaling parameter).

```{r, results="asis", echo=FALSE}
lsq = function(tau, M, s, N)
{
  start = c(mu_A = 0.3, sigma_A = 2, mu_B = 1 , sigma_B = 5, mu_M = 200)
  optim(start, fn = gof,
    tau = tau, m = M, s = s, n = N, hessian = TRUE, control = list(fnscale = 1))
}

mathout(lsq)

# Parameters' estimates
fit <-  lsq(tau, m, s, n)

est = fit$par
se = sqrt(diag(solve(fit$hessian)))
cilo = qnorm(0.025, est, se)
ciup = qnorm(0.975, est, se)

# Lower bound
#cilo

# Upper bound
#ciup

```

### Results

The calculated Goodness of Fit value is $X^2= 30.54$, $p= 0.0002$. Additionally, we proceeded with the parameters' estimation with the Least Squares Method.

The values of the estimated parameters reported in the table below are consistent with those reported by Schwarz [-@schwarz1994].

Furthermore, we can observe from the values reported in the Table below that, with few exceptions (e.i. $\tau= 100$) Miller's [-@miller1986] observed values are quite similar to the one predicted by Schwarz' [-@schwarz1994] model. Showing that, despite the calculated Goodness of Fit measure, the model proposed by Schwarz [-@schwarz1994] sufficiently fit the data.

>>>>

```{r custom-operators, echo=FALSE}
op1 <- list(
  "A\\ %*%\\ B"=quote(A %*% B),
  "A\\ %.%\\ B"=quote(A %.% B),
  "A\\ %x%\\ B"=quote(A %x% B),
  "A\\ %/%\\ B"=quote(A %/% B),
  "A\\ %%\\ B"=quote(A %% B),
  "A\\ &\\ B"=quote(A & B),
  "A\\ |\\ B"=quote(A | B),
  "xor(A,\\ B)"=quote(xor(A, B)),
  "!A"=quote(!A),
  "A\\ ==\\ B"=quote(A == B),
  "A\\ <-\\ B"=quote(A <- B))

@@

op1 <- names(op1)
if(knitr::is_latex_output())
  op1 <- sapply(op1, FUN=knitr:::escape_latex)
if(knitr::is_html_output())
  op1 <- sapply(op1, FUN=knitr:::escape_html)

op2 <- list(
  "A\\ !=\\ B"=quote(A != B),
  "A\\ ~ B"=quote(A ~ B),
  "A\\ %~~%\\ B"=quote(A %~~% B),
  "A\\ %==%\\ B"=quote(A %==% B),
  "A\\ %=~%\\ B"=quote(A %=~% B),
  "A\\ %prop%\\ B"=quote(A %prop% B),
  "A\\ %in%\\ B"=quote(A %in% B),
  "intersect(A,\\ B)"=quote(intersect(A, B)),
  "union(A,\\ B)"=quote(union(A, B)),
  "crossprod(A,\\ B)"=quote(crossprod(A, B)),
  "is.null(A)"=quote(is.null(A)))

@@

op2 <- names(op2)
if(knitr::is_latex_output())
  op2 <- sapply(op2, FUN=knitr:::escape_latex)
if(knitr::is_html_output())
  op2 <- sapply(op2, FUN=knitr:::escape_html)

op3 <- list(
  "A\\ %<->%\\ B"=quote(A %<->% B),
  "A\\ %->%\\ B"=quote(A %->% B),
  "A\\ %<-%\\ B"=quote(A %<-% B),
  "A\\ %up%\\ B"=quote(A %up% B),
  "A\\ %down%\\ B"=quote(A %down% B),
  "A\\ %<=>%\\ B"=quote(A %<=>% B),
  "A\\ %=>%\\ B"=quote(A %=>% B),
  "A\\ %<=%\\ B"=quote(A %<=% B),
  "A\\ %dblup%\\ B"=quote(A %dblup% B),
  "A\\ %dbldown%\\ B"=quote(A %dbldown% B),
  " "="")

@@

op3 <- names(names(op3))
if(knitr::is_latex_output())
  op3 <- sapply(op3, FUN=knitr:::escape_latex)
if(knitr::is_html_output())
  op3 <- sapply(op3, FUN=knitr:::escape_html)

@@

```{r, results='asis'}
zv <- Vectorize(z, vectorize.args = c('tau', 'm', 's', 'n'))

@@

The degrees of freedom result from the 13 predicted mean response times, 
minus 5 free parameters
\(\theta = \langle\mu_{\mathrm A}, \sigma_{\mathrm A}, \mu_{\mathrm B}, \sigma_{\mathrm B}, \mu_{\mathrm M}\rangle\)
(\(c\) is a redundant scaling parameter).

```{r, results="asis", echo=FALSE}
lsq = function(tau, m, s, n)
{
  invisible(theta <- c(mu_A=0.53, sigma_A=4.3, mu_B=1.34, sigma_B=11.7, mu_M=160))
  optim(theta, gof, tau=tau, m=m, s=s, n=n, hessian=TRUE)
}

mathout(call("=", quote(hat(theta)), lsq))
```

The best fitting parameter values and their confidence intervals are given in
Table \@ref(tab:params) below.

```{r params, echo=FALSE}
fit <- lsq(tau, m, s, n)
theta <- fit$par
se = sqrt(diag(solve(fit$hessian)))
cilo = qnorm(0.025, theta, se)
ciup = qnorm(0.975, theta, se)

ntheta <- lapply(lapply(FUN=as.symbol, names(theta)), FUN=mathout, flags=list(cat=FALSE))
ttheta <- lapply(theta, FUN=mathout, flags=list(cat=FALSE))
tcilo <- lapply(cilo, FUN=mathout, flags=list(cat=FALSE))
tciup <- lapply(ciup, FUN=mathout, flags=list(cat=FALSE))

t <- cbind(Parameter=ntheta, Estimate=ttheta, Lo=tcilo, Up=tciup)
knitr::kable(t, caption="Model fit", row.names=FALSE, escape=FALSE)
```

The statistic goodness-of-fit value is \(X^2(8) = 30.54\), \(p = .0002\), 
indicating some lack of fit. Given the large trial numbers in the original
study, some lack of fit is expected. For more detail, especially on fitting the
observed standard deviations, the reader is referred to the original
paper [@schwarz1994].
